---
title: "EDPS 845 Assignment 1"
author: "Tony Albano"
---

## Overview

Demonstrate your mastery of concepts and procedures covered in *An Introduction to R*, chapters 1 through 9. See Appendix A for examples.

Complete this assignment by responding to each question below using your own original code, in addition to answering any questions. Follow Google's [R coding conventions](https://google.github.io/styleguide/Rguide.xml), and use [R markdown](http://rmarkdown.rstudio.com/index.html) formatting for your responses, with example R code written as

```{r}
# Comment within some example code
```

Your file should have an Rmd extension, and you should check that it compiles to HTML using the *Knit* command in RStudio (this runs `knitr::knit()` on the file). Note that you can already *Knit* this file to see how it works.

## Exercises

### Chapter 1

1. Show that you can access the home page for the HTML R help menu, as well as help files for individual functions.
```{r}
#the HTML help menu can be accessed with:
help.start()
#to access help files for individual functions use:
?function_name
```
2. Demonstrate correct and incorrect naming of R objects.
```{r}
#correct object naming examples:
objectname
object_name

#incorrect object naming examples:
object name
01_objectname
```
3. How do you list all available objects in the R workspace? And how do you remove individual objects?
```{r}
#list all available objects in workspace
ls()

#remove individual objects
rm (object_name, object_name2)
```
### Chapter 2

1. Create and index a vector using a logical, integer, negative integer, and character vector.x
```{r}
#logical vector
y <- x[!is.na(x)]
#postive integer vector
x[1:3]
#negative integer vector
y <- x[-(4:5)]
#character vector
veggies <- c(1, 2, 3)
names(veggies) <-c("carrot", "cucumber", "broccoli")
```
2. Create a simple data frame containing a numeric variable, character variable, and factor with at least 3 levels. Use the data frame to demonstrate how R codes and handles missing data. 
```{r}
#data frame
a = c(1, 4, 8)
b = c("up", "down", "left")
c = c(TRUE, FALSE, TRUE)
direction <- factor(c)
df= data.frame(a, b, c)

#missing data
is.na(df)
         a     b     c
[1,] FALSE FALSE FALSE
[2,] FALSE FALSE FALSE
[3,] FALSE FALSE FALSE
```
### Chapter 3

1. Demonstrate how to check and change the modes and attributes of R objects.
```{r}
#check mode
mode(a)
#change mode
d <- as.character(a)
```
### Chapter 4

1. Complete the exercise on page 17, where you calculate 95% confidence intervals for state mean incomes.
```{r}
#exercise 17 for confidence intervals
state <- c("tas", "sa", "qld", "nsw", "nsw", "nt", "wa", "wa", "qld", "vic", "nsw", "vic", "qld", "qld", "sa", "tas", "sa", "nt", "wa", "vic", "qld", "nsw", "nsw", "wa", "sa", "act", "nsw", "vic", "vic", "act")
statef <- factor(state)
incomes <- c(60, 49, 40, 61, 64, 60, 59, 54, 62, 69, 70, 42, 56, 61, 61, 61, 58, 51, 48, 65, 49, 49, 41, 48, 52, 46, 59, 46, 58, 43)
incmeans <- tapply(incomes, statef, mean)
stderr <- function(x) sqrt(var(x)/length(x))
inster <- tapply(incomes, statef, stderr)
n <- table(statef)
conf <- data.frame(mean = incmeans, stderr = incster,state = names(incmeans),lower = qt(.025, n - 1), upper = qt(.025, n - 1, lower.tail = FALSE))
conf
```
2. In what order are the levels of ordered factors stored in?
```{r}
#levels are ordered alphabetically by default, or in the order they were specified explicitly.
```
### Chapter 5

1. Create an array with dimensions `c(3, 6, 2)` and use it to demonstrate array indexing.
```{r}
#create an array
array1 <- array(1:36, dim=c(3, 6, 2))
#array indexing
array1[2,,]
```
2. Create a two dimensional array of 100 integers sampled with replacement from the vector 1:20, and use an index array to set the diagonal values to 0.
```{r}
#dimensional array of 100 integers
array2 <- sample(1:20, 100, replace = TRUE)
dim(array2) <- c(10,10)
diag(array2) <- 0
```
3. Construct the same array using the `array()` function and using a vector with the `dim()` set. Prove they are the same.
```{r}
#same array using array()
array3data <- sample(1:20, 100, replace = TRUE)
array3 <- array(array3data, dim = c(10,10))
diag(array3) <- 0
#prove they are the same
identical (dim(array2), dim(array3))
```
4. Use your array from 5.1, and another with the same dimensions, to demonstrate how arithmetic works with arrays.
```{r}
#array multiplication
array2 * array3
```
5. Show how to clear the dimensions of an array.
```{r}
#clear dimension of array
dim(array3) = NULL
```
### Chapter 6

1. Demonstrate the different ways of subsetting a list.
```{r}
#subsetting a list
#creating the list
Lst <- list(name="Fred", wife="Mary", no.children=3, child.ages=c(4,7,9))
#components are:
Lst[[1]], Lst[[2]], etc.
Lst[[4]][1] #first entry of vector subsribed array
#naming a component of list
name$component_name
```
2. What is the difference between giving multiple lists to `list()` versus `c()`?
```{r}
#list() forms new lists from existing objects (e.g. multiple lists) and retains their original atributes.
Lst <- list(name_1=object_1, ..., name_m=object_m)
#c() joins together multiple lists into a single vector structure and all other attributes are discarded.
list.ABC <- c(list.A, list.B, list.C)
```
3. How do data frames in R relate to lists?
```{r}
#a data frame is a list with class "data.frame" and with certain restrictions.
```
4. What is the purpose of `attach()`?
```{r}
#attach() makes the components of a list or data fram temporarily available as variables under thier component name. This saves time by not having to quote the list name each time it is used.
```
### Chapter 7

1. Demonstrate the different approaches to reading in data from an external file.
```{r}
#read in data frame
read.table()
dataset1 <- read.table("dataset1", header=TRUE)
#scan in vector of equal length
inp <- scan("input.dat", list("",0,0))
label <- inp[[1]]; x <- inp[[2]]; y <- inp[[3]]
#access builtin datasets
data(dataset_name)
#load data from other R packages
data(package="rpart")
data(Puromycin, package="datasets")
```
### Chapter 8

1. Demonstrate the use of distribution functions to simulate data and calculate probabilities and quantiles.
```{r}
#simulate data with 500 points, mean of 50, sd of 5
simdata <- rnorm(500, 50, 5)
#check probablity for 2 standard deviations normal dist
pnorm(2.58, 0, 1)
#check quantiles for 95%
qnorm(.95, 50, 5)
```
2. Simulate data from a t distribution and check their normality.
```{r}
#simulate data from a t distribution
simdatat <- rt(500, df = n-1)
#check normality
summary(simdatat)
stem(simdatat)
hist(simdatat)
qqnorm(simdatat);qqline(simdatat)
```
### Chapter 9

1. What is the difference between `if() else` and `ifelse()`?
```{r}
#the if()else function evaluates the first element and then completes the expression based on it. The ifelse() function is a vectorized version and returns a vector of the length of its longest arguement.
```
2. Demonstrate the use of a `for()` loop.
```{r}
#create vector of day of week
day <- c(1, 2, 3, 4, 5, 6, 7)
#loop to print day of week
for (day in c(1:7)){print(paste("The day is", day))}
```